---
title: "StatsBomb"
author: "Kevin O'Donnell"
date: "2024-01-28"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries
```{r}
library(tidyverse)
library(StatsBombR)
library(SBpitch)
library(ggrepel)
```

# Loading StatsBomb EPL Event Data
```{r}
#EPL 2015-16 only
prem_comps <- FreeCompetitions() %>% 
  filter(competition_name == "Premier League",
         season_id == 27)
prem_free <- FreeMatches(prem_comps)

prem_event2 <- free_allevents(MatchesDF = prem_free, Parallel = T) 
clean_prem_event <- allclean(prem_event2)
```


# Filtering for Aerial Duels
```{r}
#16,845 aerials won - correspond to the following columns being true
#even 50/50 split in data
clean_prem_event %>%
  select(type.name, clearance.aerial_won, pass.aerial_won, shot.aerial_won, miscontrol.aerial_won) %>%
  filter(!is.na(clearance.aerial_won) | !is.na(pass.aerial_won) | !is.na(shot.aerial_won) | !is.na(miscontrol.aerial_won))

```

# Quantifying Aerial Duel Success and Success Rate
```{r}
#import standings
standings <- read.csv("standings.csv")

#changing Bournemouth's name
standings <- standings %>%
  mutate(Team = ifelse(Team=="Bournemouth", "AFC Bournemouth", Team))

#filtering for aerial duel events
a_duels <- clean_prem_event %>% 
  filter(duel.type.name=="Aerial Lost" | !is.na(clearance.aerial_won) | !is.na(pass.aerial_won) | !is.na(shot.aerial_won) | !is.na(miscontrol.aerial_won)) %>% 
  select(id:position.name, duel.type.id, duel.type.name, duel.outcome.id, duel.outcome.name, clearance.aerial_won, miscontrol.aerial_won, shot.aerial_won, pass.aerial_won, location, location.x, location.y, match_id)

#filtering only for passess off of aerial duels
passes <- clean_prem_event %>% filter(pass.aerial_won==T)

#creating result column 
a_duels <- a_duels %>%
  mutate(clearance.aerial_won = ifelse(clearance.aerial_won==T, "Clearance", clearance.aerial_won),
         pass.aerial_won = ifelse(pass.aerial_won==T, "Pass", pass.aerial_won),
         shot.aerial_won = ifelse(shot.aerial_won==T, "Shot", shot.aerial_won),
         miscontrol.aerial_won = ifelse(miscontrol.aerial_won==T, "Miscontrol", miscontrol.aerial_won))

a_duels <- a_duels %>%
  mutate(result = coalesce(duel.type.name, clearance.aerial_won, pass.aerial_won, shot.aerial_won, miscontrol.aerial_won))

#creating duel win column (1 if won, 0 if lost)
a_duels <- a_duels %>%
  mutate(W = (result !="Aerial Lost"))

# Overall breakdown of result types
a_duels %>%
  group_by(result) %>%
  summarise(pct = n()/nrow(a_duels), n=n())

# breakdown of result types by team
a_duels %>%
  group_by(team.name, result) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  group_by(team.name) %>%
  mutate(n2=sum(n), pct=n/n2) %>%
  arrange(team.name, -pct)

# breakdown of W vs L by team
a_duels %>%
  group_by(team.name, W = (result !="Aerial Lost")) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  group_by(team.name) %>%
  mutate(n2=sum(n), pct=n/n2) %>%
  arrange(-pct) %>%
  filter(W==T) %>%
  select(-W) 

# breakdown of result types by player (min 45 duels - Q1)
a_duels %>%
  group_by(player.name, result) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  group_by(player.name) %>%
  mutate(n2=sum(n), pct=n/n2) %>%
  arrange(result, -pct) %>%
  filter(n2>45)

# breakdown of W vs L by player (min 45 duels - Q1)
a_duels %>%
  group_by(player.name, W = (result !="Aerial Lost")) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  group_by(player.name) %>%
  mutate(n2=sum(n), pct=n/n2) %>%
  arrange(-pct) %>%
  filter(W==T, n2>45) 
```

# Visualizing Duel Success Spatially
```{r}
#note: right side is attacking, left is defending

#heatmap - splitting into 24 even zones
heatmap <- a_duels
heatmap$xbin <- cut(heatmap$location.x, breaks = seq(from=0, to=120, by = 20),include.lowest=TRUE )
heatmap$ybin <- cut(heatmap$location.y, breaks = seq(from=0, to=80, by = 20),include.lowest=TRUE) 

#plotting mean win rate of duels for each team compared to average (pure wins and losses, not long term)
heatmap = heatmap %>%
  group_by(team.name) %>%
  mutate(success_rate = mean(W),
       n = n()) %>%
  group_by(team.name, xbin, ybin) %>%
  summarise(success_rate = max(success_rate),
          n=max(n),
          bin_duels_success_rate = mean(W),
          bin_duel_n = n(),
          location.x = median(location.x),
          location.y = median(location.y)) %>%
  group_by(xbin, ybin) %>%
  mutate(league_ave = mean(bin_duels_success_rate)) %>%
  group_by(team.name, xbin, ybin) %>%
  mutate(diff_vs_ave = bin_duels_success_rate - league_ave)


library(grid)
defensiveactivitycolors <- c("#dc2429", "#dc2329", "#df272d", "#df3238", "#e14348", "#e44d51",
"#e35256", "#e76266", "#e9777b", "#ec8589", "#ec898d", "#ef9195",
"#ef9ea1", "#f0a6a9", "#f2abae", "#f4b9bc", "#f8d1d2", "#f9e0e2",
"#f7e1e3", "#f5e2e4", "#d4d5d8", "#d1d3d8", "#cdd2d6", "#c8cdd3", "#c0c7cd",
"#b9c0c8", "#b5bcc3", "#909ba5", "#8f9aa5", "#818c98", "#798590",
"#697785", "#526173", "#435367", "#3a4b60", "#2e4257", "#1d3048",
"#11263e", "#11273e", "#0d233a", "#020c16") 

ggplot(data= heatmap %>% filter(bin_duel_n>10, ), aes(x = location.x, y = location.y, fill = diff_vs_ave, group =diff_vs_ave)) +
 geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) + #2
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "white", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 annotate("point", x = 12 , y = 40, colour = "white", size = 1.05) +
 annotate("point", x = 108 , y = 40, colour = "white", size = 1.05) +
 annotate("path", colour = "white", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 annotate("point", x = 60 , y = 40, colour = "white", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
 annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=8,family="Arial", hjust=0.5, vjust=0.5),
 plot.subtitle = element_text(size = 13, family="Arial", hjust = 0.5),
 axis.text.y=element_blank(),
 legend.title = element_blank(),
 legend.text=element_text(size=8,family="Arial"),
 legend.key.size = unit(.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 15,
family="Arial", colour = "black", hjust = 0.5),
 legend.direction = "vertical",
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=7,family="Arial")) + #4
 scale_y_reverse() + #5
 scale_fill_gradientn(colours = defensiveactivitycolors, trans = "reverse", labels =
scales::percent_format(accuracy = 1), limits = c(0.20, -0.20), breaks = scales::breaks_extended(n = 9)) + #6
 labs(title = "Duel Win Rate vs League Average", subtitle = "EPL, 2015/16") + #7
 coord_fixed(ratio = 95/100) +
 annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last",
 length=unit(2.55,"mm")), gp=gpar(col="black", fill=NA, lwd=2.2)),
 xmin=25, xmax = 95, ymin = -83, ymax = -83) + #9
 facet_wrap(~team.name)+ #10
 guides(fill = guide_legend(reverse = TRUE)) 
```

# Defining Longer Term Duel Success - Did it lead to above average possession length, shot, etc.
```{r}
#distribution of possession time
clean_prem_event %>%
  group_by(match_id, possession) %>%
  slice(n()) %>%
  ggplot(aes(x=TimeInPoss)) +
  geom_boxplot()

summary((clean_prem_event %>%
  group_by(match_id, possession) %>%
  slice(n()))$TimeInPoss)
#mean = 23.329
#median = 11.708

#filtering event data to only passes (specific type of win that would extend possession)
#also including the event directly following passses to ensure that the same possession is captured
one_after <- clean_prem_event %>% 
  filter((!is.na(pass.aerial_won)) | (!is.na(lag(pass.aerial_won,1)))) %>% 
  select(id:position.name, duel.type.id, duel.type.name, duel.outcome.id, duel.outcome.name, clearance.aerial_won, miscontrol.aerial_won, shot.aerial_won, pass.aerial_won, location, location.x, location.y, match_id) %>%
  mutate(match_pos = paste(match_id, possession))

#pulling final play of ensuing possession following duel win
post_duels <- clean_prem_event %>% 
  mutate(match_pos = paste(match_id, possession)) %>%
  filter(match_pos %in% unique(one_after$match_pos)) %>%
  group_by(match_id, possession) %>%
  slice(n())

post_duels %>% 
  group_by(type.name) %>%
  summarise(n=n()) %>%
  arrange(-n)
#successful endings: possession > average possession time, foul won, goal keeper, shot

# now get possession time for those possessions, join with duels
a_duels <- a_duels %>%  mutate(match_pos = paste(match_id, possession))

duel_and_post <- post_duels %>%
  select(match_pos, TimeInPoss, type.name.end = type.name) %>%
  right_join(a_duels, by='match_pos')

#creating long-term success variable/metric
duel_and_post <- duel_and_post %>%
  mutate(success = ifelse(W==T & (TimeInPoss>23.329 |
                            result=="Shot" |
                            type.name.end %in% c("Foul Won", "Goal Keeper", "Shot")), 1, 0)) %>%
  mutate(success = ifelse(is.na(success), 0, success)) %>%
  mutate(end_in_shot = ifelse(W==T & 
                            (result=="Shot" |
                            type.name.end %in% c("Foul Won", "Goal Keeper", "Shot")), 1, 0)) %>%
  mutate(end_in_shot = ifelse(is.na(end_in_shot), 0, end_in_shot))

#breakdown of success to failure for LDSR
duel_and_post %>% group_by(success) %>%
  summarise(n())
```

# Comparing and Visualizing Duel Win Rate + Long Term Duel Success
```{r}
#team success rate and shot rate (possession ends in shot)
success_shot_rate_team <- duel_and_post %>%
  group_by(team.name) %>%
  filter(W==T) %>%
  summarise(success_rate = mean(success), shot_rate = mean(end_in_shot)) %>%
  arrange(-success_rate)

#joining team success rates with standings
success_shot_rate_team <- left_join(success_shot_rate_team, standings, by=c('team.name'='Team'))

#correlation between success rate and points in the table
cor(success_shot_rate_team$success_rate, success_shot_rate_team$Pts)

#saving win rate by team
win_rate_team <- a_duels %>%
  group_by(team.name, W = (result !="Aerial Lost")) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  group_by(team.name) %>%
  mutate(n2=sum(n), pct=n/n2) %>%
  arrange(-pct) %>%
  filter(W==T) 

#joining team table with win rate
success_shot_rate_team <- left_join(success_shot_rate_team, win_rate_team, by=c('team.name'))

#correlation between success rate and win rate
cor(success_shot_rate_team$success_rate, success_shot_rate_team$pct)
cor(filter(success_shot_rate_team, !team.name %in% c("Aston Villa", "Leicester City"))$success_rate, filter(success_shot_rate_team, !team.name %in% c("Aston Villa", "Leicester City"))$Pts)

#correlation between win rate and points in table
cor(success_shot_rate_team$pct, success_shot_rate_team$Pts)
cor(filter(success_shot_rate_team, !team.name %in% c("Aston Villa", "Leicester City"))$pct, filter(success_shot_rate_team, !team.name %in% c("Aston Villa", "Leicester City"))$Pts)

#plotting duel win rate vs points
ggplot(success_shot_rate_team, aes(Pts, pct)) +
  geom_point() +
  geom_smooth(method = "lm", color = "#dc2429", se = FALSE, linetype="dashed", alpha=0.6) +
  theme(
    axis.text.y = element_text(size = 11, color = "#333333", family = "Arial"),
    axis.title = element_text(size = 12, color = "#333333", family = "Arial"),
    axis.text.x = element_text(size = 11, color = "#333333", family = "Arial"),
    axis.ticks = element_line(),
    panel.background = element_rect(fill = "white", colour = "white"),
    plot.background = element_rect(fill = "white", colour = "white"),
    panel.grid.major = element_line(color = "gray", linetype = "dashed"),  # Add gridlines
    panel.grid.minor = element_line(color = "gray", linetype = "dashed"),  # Add gridlines
    plot.title = element_text(size = 16, color = "#333333", family = "Arial", face = "bold"),
    plot.subtitle = element_text(size = 14, color = "#333333", family = "Arial", face = "bold"),
    plot.caption = element_text(color = "#333333", family = "Arial", size = 10),
    text = element_text(family = "Arial"),
    legend.title = element_blank()
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Aerial Duel Win Rate vs Points", subtitle = "EPL, 2015/16", x = "Points", y = "Duel Win Rate") +
  geom_label_repel(aes(label = team.name), data = success_shot_rate_team %>% filter(team.name %in% c("Aston Villa", "Leicester City")), box.padding = 0.5)


#plotting long term success rate vs points
ggplot(success_shot_rate_team, aes(Pts, success_rate)) +
  geom_point() +
  geom_smooth(method = "lm", color = "#dc2429", se = FALSE, linetype="dashed", alpha=0.6) +
  theme(
    axis.text.y = element_text(size = 11, color = "#333333", family = "Arial"),
    axis.title = element_text(size = 12, color = "#333333", family = "Arial"),
    axis.text.x = element_text(size = 11, color = "#333333", family = "Arial"),
    axis.ticks = element_line(),
    panel.background = element_rect(fill = "white", colour = "white"),
    plot.background = element_rect(fill = "white", colour = "white"),
    panel.grid.major = element_line(color = "gray", linetype = "dashed"),  # Add gridlines
    panel.grid.minor = element_line(color = "gray", linetype = "dashed"),  # Add gridlines
    plot.title = element_text(size = 16, color = "#333333", family = "Arial", face = "bold"),
    plot.subtitle = element_text(size = 14, color = "#333333", family = "Arial", face = "bold"),
    plot.caption = element_text(color = "#333333", family = "Arial", size = 10),
    text = element_text(family = "Arial"),
    legend.title = element_blank()
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Long-Term Duel Success Rate vs Points", subtitle = "EPL, 2015/16", x = "Points", y = "Long-Term Duel Success Rate (LDSR)") +
  geom_label_repel(aes(label = team.name), data = success_shot_rate_team %>% filter(team.name %in% c("Aston Villa", "Leicester City", "Arsenal", "Newcastle United")), box.padding = 0.7)


#player success rates
#note: LDSR is more of a team metric because it is dependent on how the team capitalizes on aerial wins, not just how often individual players win their duels. This metric is computed for players for exploratory purposes 
success_shot_rate_player <- duel_and_post %>%
  group_by(player.name) %>%
  #filter(W==T) %>%
  summarise(success_rate = mean(success), shot_rate = mean(end_in_shot), n=n(), position.id) %>%
  arrange(-success_rate) %>%
  mutate(position_group = case_when(position.id<=8 ~ "Defender",
                                    position.id==17 | position.id>=21 ~ "Attacker",
                                    .default = "Midfielder")) %>%
  select(-position.id) %>%
  unique()

#most successful players above 45 duels attempted
success_shot_rate_player %>%
  filter(n>45) %>%
  arrange(-success_rate)

#saving win rate by player
win_rate_player <- a_duels %>%
  group_by(player.name, W = (result !="Aerial Lost")) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  group_by(player.name) %>%
  mutate(n2=sum(n), pct=n/n2) %>%
  arrange(-pct) %>%
  filter(W==T) 

#joining player table with win rate
success_shot_rate_player <- left_join(success_shot_rate_player, win_rate_player, by='player.name')

#correlation between win rate and success rate on player level 
cor(success_shot_rate_player$pct, success_shot_rate_player$success_rate, use = "complete.obs")

cor(filter(success_shot_rate_player, n.x>45)$pct, filter(success_shot_rate_player, n.x>45)$success_rate, use = "complete.obs")
cor(filter(success_shot_rate_player, n.x>45)$pct, filter(success_shot_rate_player, n.x>45)$shot_rate, use = "complete.obs")

#plot of win rate and success rate on player level 
ggplot(filter(success_shot_rate_player, n.x>45), aes(x=pct, y=success_rate)) +
  geom_point() 

#success rates and win rates by position group (attacker, defender, midfielder)
success_shot_rate_player %>%
  group_by(position_group) %>%
  summarise(mean(success_rate), mean(shot_rate), mean(n.x), mean(pct, na.rm=T))
```

# Visualizing Long Term Duel Success Spatially
```{r}
#heatmap 
heatmap <- duel_and_post
heatmap$xbin <- cut(heatmap$location.x, breaks = seq(from=0, to=120, by = 20),include.lowest=TRUE )
heatmap$ybin <- cut(heatmap$location.y, breaks = seq(from=0, to=80, by = 20),include.lowest=TRUE) 

#plotting mean success rate of duels for each team compared to average
heatmap = heatmap %>%
  group_by(team.name) %>%
  mutate(success_rate = mean(success),
       n = n()) %>%
  group_by(team.name, xbin, ybin) %>%
  summarise(success_rate = max(success_rate),
          n=max(n),
          bin_duels_success_rate = mean(success),
          bin_duel_n = n(),
          location.x = median(location.x),
          location.y = median(location.y)) %>%
  group_by(xbin, ybin) %>%
  mutate(league_ave = mean(bin_duels_success_rate)) %>%
  group_by(team.name, xbin, ybin) %>%
  mutate(diff_vs_ave = bin_duels_success_rate - league_ave)


library(grid)
defensiveactivitycolors <- c("#dc2429", "#dc2329", "#df272d", "#df3238", "#e14348", "#e44d51",
"#e35256", "#e76266", "#e9777b", "#ec8589", "#ec898d", "#ef9195",
"#ef9ea1", "#f0a6a9", "#f2abae", "#f4b9bc", "#f8d1d2", "#f9e0e2",
"#f7e1e3", "#f5e2e4", "#d4d5d8", "#d1d3d8", "#cdd2d6", "#c8cdd3", "#c0c7cd",
"#b9c0c8", "#b5bcc3", "#909ba5", "#8f9aa5", "#818c98", "#798590",
"#697785", "#526173", "#435367", "#3a4b60", "#2e4257", "#1d3048",
"#11263e", "#11273e", "#0d233a", "#020c16") 

ggplot(data= heatmap %>% filter(bin_duel_n>10, ), aes(x = location.x, y = location.y, fill = diff_vs_ave, group =diff_vs_ave)) +
 geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) + #2
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "white", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 annotate("point", x = 12 , y = 40, colour = "white", size = 1.05) +
 annotate("point", x = 108 , y = 40, colour = "white", size = 1.05) +
 annotate("path", colour = "white", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 annotate("point", x = 60 , y = 40, colour = "white", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
 annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=8,family="Arial", hjust=0.5, vjust=0.5),
 plot.subtitle = element_text(size = 13, family="Arial", hjust = 0.5),
 axis.text.y=element_blank(),
 legend.title = element_blank(),
 legend.text=element_text(size=8,family="Arial"),
 legend.key.size = unit(.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 15,
family="Arial", colour = "black", hjust = 0.5),
 legend.direction = "vertical",
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=7,family="Arial")) + #4
 scale_y_reverse() + #5
 scale_fill_gradientn(colours = defensiveactivitycolors, trans = "reverse", labels =
scales::percent_format(accuracy = 1), limits = c(0.20, -0.20), breaks = scales::breaks_extended(n = 9)) + #6
 labs(title = "Long Term Duel Success Rate vs League Average", subtitle = "EPL, 2015/16") + #7
 coord_fixed(ratio = 95/100) +
 annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last",
 length=unit(2.55,"mm")), gp=gpar(col="black", fill=NA, lwd=2.2)),
 xmin=25, xmax = 95, ymin = -83, ymax = -83) + #9
 facet_wrap(~team.name)+ #10
 guides(fill = guide_legend(reverse = TRUE)) 
```


# Correlation of Long Term Success Rate in Each Zone with Winning (Pts)
```{r}
heatmap <- duel_and_post
heatmap$xbin <- cut(heatmap$location.x, breaks = seq(from=0, to=120, by = 20),include.lowest=TRUE )
heatmap$ybin <- cut(heatmap$location.y, breaks = seq(from=0, to=80, by = 20),include.lowest=TRUE) 

#plotting correlation of success rate to points across the field (split into thirds)
heatmap = heatmap %>%
  group_by(team.name) %>%
  mutate(success_rate = mean(success),
       n = n()) %>%
  group_by(team.name, xbin, ybin) %>%
  summarise(success_rate = max(success_rate),
          n=max(n),
          bin_duels_success_rate = mean(success),
          bin_duel_n = n(),
          location.x = median(location.x),
          location.y = median(location.y)) %>%
  group_by(xbin, ybin) %>%
  mutate(league_ave = mean(bin_duels_success_rate)) %>%
  group_by(team.name, xbin, ybin) %>%
  mutate(diff_vs_ave = bin_duels_success_rate - league_ave)

corr_map <- heatmap %>%
  left_join(standings, by=c('team.name'='Team')) %>%
  group_by(xbin, ybin) %>%
  summarise(correlation = cor(diff_vs_ave, Pts, use = "complete.obs"), location.x =median(location.x), location.y =median(location.y)) %>%
  arrange(-correlation)


ggplot(data= corr_map, aes(x = location.x, y = location.y, fill = correlation, group =correlation)) +
 geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) + #2
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "white", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 annotate("point", x = 12 , y = 40, colour = "white", size = 1.05) +
 annotate("point", x = 108 , y = 40, colour = "white", size = 1.05) +
 annotate("path", colour = "white", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 annotate("point", x = 60 , y = 40, colour = "white", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
 annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=8,family="Arial", hjust=0.5, vjust=0.5),
 plot.subtitle = element_text(size = 10, family="Arial", hjust = 0.5),
 axis.text.y=element_blank(),
 legend.title = element_blank(),
 legend.text=element_text(size=8,family="Arial"),
 legend.key.size = unit(.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 12,
family="Arial", colour = "black", hjust = 0.5),
 legend.direction = "vertical",
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=7,family="Arial")) + #4
 scale_y_reverse() + #5
 scale_fill_gradientn(colours = defensiveactivitycolors, trans = "reverse", labels =
scales::percent_format(accuracy = 1), limits = c(0.50, -0.50), breaks = scales::breaks_extended(n = 9)) + #6
 labs(title = "Correlation between Long Term Duel Success Rate and Cumulative Points", subtitle = "EPL Teams, 2015/16") + #7
 coord_fixed(ratio = 95/100) +
 annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last",
 length=unit(2.55,"mm")), gp=gpar(col="black", fill=NA, lwd=2.2)),
 xmin=25, xmax = 95, ymin = -83, ymax = -83) + #9
 guides(fill = guide_legend(reverse = TRUE)) 
```


# Correlation of Long Term Success Rate in Each Zone with Winning (Pts)
```{r}
heatmap <- duel_and_post
heatmap$xbin <- cut(heatmap$location.x, breaks = seq(from=0, to=120, by = 20),include.lowest=TRUE )
heatmap$ybin <- cut(heatmap$location.y, breaks = seq(from=0, to=80, by = 20),include.lowest=TRUE) 

#plotting correlation of long-term success rate in each zone with points
heatmap = heatmap %>%
  group_by(xbin, ybin) %>%
  summarise(success = sum(success),
       n = n(),
       location.x = median(location.x),
      location.y = median(location.y)) %>%
  mutate(failure = n-success,
         pct = success/n)


ggplot(data= heatmap, aes(x = location.x, y = location.y, fill = pct, group =pct)) +
 geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) + #2
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "white", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 annotate("point", x = 12 , y = 40, colour = "white", size = 1.05) +
 annotate("point", x = 108 , y = 40, colour = "white", size = 1.05) +
 annotate("path", colour = "white", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 annotate("point", x = 60 , y = 40, colour = "white", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
 annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=8,family="Arial", hjust=0.5, vjust=0.5),
 plot.subtitle = element_text(size = 10, family="Arial", hjust = 0.5),
 axis.text.y=element_blank(),
 legend.title = element_blank(),
 legend.text=element_text(size=8,family="Arial"),
 legend.key.size = unit(.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 12,
family="Arial", colour = "black", hjust = 0.5),
 legend.direction = "vertical",
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=7,family="Arial")) + #4
 scale_y_reverse() + #5
 scale_fill_gradientn(colours = defensiveactivitycolors, trans = "reverse", labels =
scales::percent_format(accuracy = 1), limits = c(.4, 0), breaks = scales::breaks_extended(n = 9)) + #6
 labs(title = "Long Term Duel Success Rate by Zone", subtitle = "EPL League Average, 2015/16") + #7
 coord_fixed(ratio = 95/100) +
 annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last",
 length=unit(2.55,"mm")), gp=gpar(col="black", fill=NA, lwd=2.2)),
 xmin=25, xmax = 95, ymin = -83, ymax = -83) + #9
 guides(fill = guide_legend(reverse = TRUE)) 
```




